Apparent preferences for cognitive effort fade when multiple forms of effort and delay are interleaved in a foraging environment

### 2020-10-27 Cost_manuscript.Rmd
two last intro pars

  Evaluating whether to pursue a desired outcome often requires assessing the demands required for obtaining it. The intuitive reaction to this evaluation is aversion. In psychology and economics, demands for both time and effort have traditionally been identified as costs that detract from the net subjective value of rewards (Westbrook & Braver, 2015). Indeed, decision makers tend to prefer low-effort and immediate rewards over larger rewards that demand longer delays and higher effort (Ainslie, 1975; Hull, 1943; Kool & Botvinick, 2018; Kool, McGuire, Rosen, & Botvinick, 2010; Treadway, Buckholtz, Schwartzman, Lambert, & Zald, 2009; Westbrook, Kester, & Braver, 2013). However, people often change their attitudes towards exerting effort in everyday decisions, resulting in pursual of effortful avenues instead (Inzlicht, Shenhav, & Olivola, 2017). Much experimental and theoretical work surrounding the sporadic attractiveness of effort suggests a reciprocal relationship between effort and outcome, whereby exerting effort boosts the perceived value of its outcomes (kacelnik), and positive outcomes imbue value into their preceding effortful behavior (learned industriousness cite). Even though the prevalent emphasis on these retrospective value associations can help explain a variety of scenarios, it is still in its infancy (be more explicit).

  What is often missing from these accounts is the common case of choosing among behavioral demands, which is partly due to the scope of experimental designs. Studies tend to compare the reduction of subjective value imposed by each type of demand, often through discounting functions that capture how much less appealing a reward becomes as a function of the amount or intensity of its associated behavior [1,3,5,8,9,14]. Although this line of work suggests that discounting functions are similar for delay and effort [8,15,16], there is ongoing debate about the shape of the discounting function across effort types [2,8,11,17,18], and discount rates do not always correlate across cost domains [9,16]. Rather than arising from changes in the decay of value, flexible preferences for effort might stem from the availability of other types of demands in the same choice environment—in other words, from its context (Kool & Botvinick, 2018). For example, exerting high effort for a reward might be aversive, but the same level of work might become valuable when compared to the boredom of passive waiting. Now add charitable giving evidence from inzlicht paper.

  Single choice foraging paradigms provide a natural way to probe preferences among demands, by asking individuals whether the current offer outweighs what could be obtained and experienced during that time instead (Mobbs, Trimmer, Blumstein, & Dayan, 2018). In addition, foraging tasks can allow individuals to emphasize different elements of the decision context, and rest on a rich modeling history that can help to computationally capture such emphases (Stephens & Krebs, 2019). Foraging has attracted high scientific interest due to its ecological validity, which is rooted in evolutionary choice behaviors [49]. The paradigm requires the decision maker to choose whether to engage with a single presented offer or skip it in search of a better one. A key advantage of foraging paradigms is that people can rapidly adopt stable, rewardmaximizing strategies [50]. Add that Fawcett bit…

  In this study, we tested the hypothesis that preferences would vary depending on whether individuals face a single or multiple types of behavioral demands (i.e. cognitive effort, physical effort, and delay). We leveraged the ecological and computational richness of foraging paradigms to provide new insights on a number of open questions. First, the existence of a well-defined optimal choice strategy in foraging designs was useful for identifying when demands boost the value of rewards (i.e. overharvesting poor offers). Second, increasing the level of exposure to types of demands over two experiments let us probe whether their cost is intrinsic or relative. Third, confronting participants with both mental and physical effort let us probe whether these preferences are domain general or task-specific. Finally, modeling helped us disambiguate what drove these preferences.

>>>>

  Evaluating whether to pursue a desired outcome often requires assessing the demands required for obtaining it. The intuitive reaction to this evaluation is to find demands aversive. In psychology and economics, demands for both time and effort have traditionally been identified as costs that detract from the net subjective value of rewards (Westbrook & Braver, 2015). Indeed, decision makers tend to prefer low-effort and immediate rewards over larger rewards that demand longer delays and higher effort (Ainslie, 1975; Hull, 1943; Kool & Botvinick, 2018; Kool, McGuire, Rosen, & Botvinick, 2010; Treadway, Buckholtz, Schwartzman, Lambert, & Zald, 2009; Westbrook, Kester, & Braver, 2013). However, people often change their attitudes towards exerting effort in everyday decisions, resulting in pursual of effortful avenues instead. Emerging experimental and theoretical work has chiefly explained this phenomenon by suggesting a reciprocal relationship between effort and outcome, whereby exerting effort boosts the perceived value of its outcomes (Kacelnik and Marsh, 2002), and positive outcomes imbue value into their preceding effortful behavior (i.e. learned industriousness; Eisenberger, 1992). Though intuitively valid and empirically supported, this explanation leaves open questions about what intrinsic features of an effortful task can make it attractive beyond its associated outcome, and how contextual factors mold their perceived costs (see Inzlicht, Shenhav, and Olivola, 2018, for a review on the attractiveness of effort).

  What is often missing from these accounts is the common case of choosing among behavioral demands, partly due to the scope of traditional experimental designs. Studies tend to compare the reduction of subjective value imposed by each type of demand, often through discounting functions that capture how much less appealing a reward becomes as a function of the amount or intensity of its associated behavior [1,3,5,8,9,14]. Although this line of work suggests that discounting functions are similar for delay and effort [8,15,16], there is ongoing debate about the shape of the discounting function across effort types [2,8,11,17,18], and discount rates do not always correlate across cost domains [9,16]. However, flexible preferences for effort could arise from the availability of other types of demands in the same choice environment—in other words, from its context (Kool & Botvinick, 2018). For example, exerting high effort for a reward might be aversive, but the same level of work might become valuable when compared to the boredom of passive waiting. Conversely, even though people tend to donate more to charitable events that require them to perform effortful activities (i.e. the martyrdom effect), they are also less likely to engage in such events once an easier alternative to raise funds becomes available (Olivola and Shafir, 2013).

  Single choice foraging paradigms provide a natural way to probe preferences among demands, by asking individuals whether the current offer outweighs what could be obtained and experienced during that time instead (Mobbs, Trimmer, Blumstein, & Dayan, 2018). In addition, foraging tasks can allow individuals to emphasize different elements of the decision context, and rest on a rich modeling history that can help to computationally capture such emphases (Stephens & Krebs, 2019). Foraging has attracted high scientific interest due to its ecological validity, which is rooted in evolutionary choice behaviors [49]. The paradigm requires the decision maker to choose whether to engage with a single presented offer or skip it in search of a better one. A key advantage of foraging paradigms is that people can rapidly adopt stable, rewardmaximizing strategies [50]. Add that Fawcett bit and modeling…

  In this study, we tested the hypothesis that preferences would vary depending on whether individuals face a single or multiple types of behavioral demands (i.e. cognitive effort, physical effort, and delay). We leveraged the ecological and computational richness of foraging paradigms to provide new insights on a number of open questions. First, the existence of a well-defined optimal choice strategy in foraging designs was useful for identifying when demands boost the value of rewards (i.e. overharvesting poor offers). Second, increasing the level of exposure to types of demands over two experiments let us probe whether their cost is intrinsic or relative. Finally, confronting participants with both mental and physical effort let us probe whether these preferences are domain general or task-specific, and in the case of the latter, isolate factors that influence preferences beyond the often cited associative relationship between effort and outcome (e.g. features of an effortful task that makes it more attractive).
